---
title: "R STA120"
author: "Zofia Baranczuk"
date: "March 13, 2024"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Data loading:
Load from whttps://user.math.uzh.ch/furrer/download/sta121/ the following data sets:

elephant.csv

Meatprice.txt

Penguins.RData 


```{r elephant}
elephant <- read.csv("./Data/elephant.csv")
View(elephant)
```

```{r meatprice}
meatprice <- read.table("./Data/Meatprice.txt", header = TRUE)
View(meatprice)

```



```{r Penguins}
load("./Data/Penguins.RData")
#a<- load("./Data/Penguins.RData")
#a
View(Penguins)
```


# 2. Working with data frames

Dataset Seatbelts is available in R from the package datasets. 
You should have access to this data set without loading.
Use ?Seatbelts to read about the dataset. 
First we will change the data set to be a data frame (not a Time series), see the code. 
```{r}
library(datasets)
Seatbelts_df <- data.frame(as.matrix(Seatbelts), date=time(Seatbelts)) 
View(Seatbelts_df)
# convert Time Series to data frame so that we can work with that. This step is not relevent for us at the moment, we just want to have a data frame to work with

```
1. **Add a column** representing the ratio of rear passengers killed to front passengers killed
```{r}
Seatbelts_df$ratio <- Seatbelts_df$rear/Seatbelts_df$front

```

2. In which month there was the lowest number of killed drivers? What was this number?
```{r}
n <- which.min(Seatbelts_df$DriversKilled)
Seatbelts_df$DriversKilled[n]
n

minDK <- min(Seatbelts_df$DriversKilled)
minDK

Seatbelts_df$Date[minDK]


which(Seatbelts_df$DriversKilled==minDK)

dateminDK <- Seatbelts_df$date[n]
dateminDK

```


3. In which months there were less than 5 van drivers killed? In how many months was that the case?
```{r}
ind <- which(Seatbelts_df$VanKilled<5)
ind
my_months <- Seatbelts_df$date[ind]
my_months

# how many months have happened 
length(my_months)

```

4. How many drivers were killed on average in one month in 1971?
```{r}
Seatbelts_df$year <- floor(Seatbelts_df$date)
sum(Seatbelts_df$DriversKilled[(Seatbelts_df$year==1971)])/12

mean(Seatbelts_df$DriversKilled[which(Seatbelts_df$year==1971)])



```
5. (*) In which year the number of killed drivers was the highest?

```{r}
## method 1
Seatbelts_df$year <- floor(Seatbelts_df$date)
drivers_yearly <- aggregate(Seatbelts_df$drivers, by=list(Seatbelts_df$year), FUN = sum) # aggregate 的默认方法是mean,此处强调是sum
#以上方法的替代方法
# drivers_yearly <- aggregate(SDriversKilled ~ year, data = Seatbelts_df, FUN = sum)


ind <- which.max(drivers_yearly$x)
drivers_yearly[ind,]

## method 2:
y_dk <- data.frame(year = c(min(Seatbelts_df$year): max(Seatbelts_df$year)))
y_dk$sum <- 0

for(y  in c(min(Seatbelts_df$year): max(Seatbelts_df$year))){
  seat_y <- Seatbelts_df[Seatbelts_df$year == y,]
  y_dk$sum[y - min(Seatbelts_df$year) + 1] <- sum(seat_y$Driverskilled)
}

y_dk$year[which.max(y_dk$sum)]
which.max(y_dk$sum)
max(y_dk$sum)
View(y_dk)

```
The highest number of deaths was in 1973, 23850 drivers died.  

6. Create a scatter plot Drivers killed against time.
Use different colors for the data points representing obligatory seat belts and non obligatory.
```{r}
plot(Seatbelts_df$date, Seatbelts_df$DriversKilled, col=Seatbelts_df$law +5)
```


7. Choose the rows in which seat belts were obligatory.
Create a data frame which contains only the rows in which the seat belts where obligatory. 
Create another data frame, which contains only the rows in which the seat belts where not obligatory.
```{r}
ind <- which(Seatbelts_df$law == TRUE)
Seatbelts_obl <- Seatbelts_df[ind,]

Seatbelts_nonobl <- Seatbelts_df[-ind,]

plot(Seatbelts_nonobl$date, Seatbelts_nonobl$DriversKilled)
plot(Seatbelts_obl$date, Seatbelts_obl$DriversKilled)

```


# 3. Loops

1. a) Plot a density of a standard normal random variable. Superimpose the density of a t-distribution with one
degree of freedom in blue color.
b) How large do you have to choose the degrees of freedom of a t-distribution such that the difference between
its density and the one of a standard normal distribution is smaller than 0.001 in every location? Hint: Use
a loop and check the difference between the densities for different degrees of freedom.

```{r}
# a)
x <- seq( -6, to=6, length=500)
plot( x, dnorm(x), type="l") 
lines( x, dt(x, df=1), col=4) 

# b)
x <- seq( 0, to=6, length=1000) # by symmetry
maxdelta <- numeric(1000)
for( i in 1:1000) 
  maxdelta[i] <- max(abs(dnorm(x) - dt(x, df=i)))

max(which( maxdelta > 0.001)) 


```
For i = 136 or more, the maximal difference between the density of the standard normal distribution and the density of the t-distribution with i degrees of freedom, is smaller than 0.001 in every location.


###########
Compute the probability that when tossing a fair coin 20 times, we will have even number of heads. 
```{r}
sum(dbinom(x=seq(0,20,by=2),size = 20, prob=0.5))


```
#######################
Let X~Poisson(2). Compute P(X in {6,7,8,9,10,11,12,13}) 
```{r}
P<- ppois(q = 13, lambda = 2)-ppois(q = 5, lambda = 2)
P
#or
sum(dpois(x = c(6:13), lambda = 2))
```
